How has the Brazilian Amazon been constructed as a problem:presidential speeches and transnational politics since 1985

### 2022-12-08 Submission/draft/draft_resubmission.rmd
Added draft folder to submission

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(readxl)
library(ggplot2)
library(gridExtra)
library(modelsummary)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

pop <-  pop %>% select(-c(title, date, location))

@@

@@

                                            location), "Amazonian Countries", location_cat),
                               levels = c("International", "Amazonian States",
                                          "Amazonian Countries", "Brasilia",
                                          "Non Amazonian States")))

```

@@

\pagebreak

# 1 Introduction

*We need to protect the Amazon from foreign interests. We need to exploit the Amazon's natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem.* Each of these statements contains an implicit assumption of what needs to be solved, or in other words, it represents the Amazon as a particular problem: national sovereignty, economic integration, social development, and environmental conservation, respectively.These specific problems touch on common understandings of the Amazon as part of the larger socio-cultural history of the country.Different governments throughout Brazilian history have been identified as proponents of a specific view of the Amazon [@drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021].

We conceptualize the Amazon region, forest, and people to be a policy object, this is a specific issue that deserves dedicated policy attention. While the military dictatorship(1964-85) is associated with understanding the Amazon as issues of national sovereignty and economic integration, Sarney's (1985-1990) and Lula's (2003-2010)presidencies are often tied to environmental conservation. These monolithic representations advance the view that specific governments understand and act towards the Amazon as one problem, hiding within government diversity of both policies and perspectives. Albeit the current calls to understand the environment as a social-cultural construction and to identify the effect of culture on environmental outcomes [@lepolaindewaroux2021], we lack empirical accounts of how the Brazilian Amazon has been constructed as a problem in discourses over time, by geographical location, and between or within, governments. Building on @hirschman1963 conceptualization of chosen problems and @bacchi2009 theory of problem-representation in policy, we propose a framework for identifying problem-constructions in discourse and investigate how the Brazilian Amazon has been constructed as a problem in transnational political discourses. We understand transnationalism[^1] as relationships that transcend nation-states, from the local level to the supranational level encompassing non-state actors [@keck1998].

[^1]: We use the noun transnationalism, the adverb transnationally, and the adjective transnational interchangeably

Although problem-construction takes place in a series of instances, we analyze the case of speeches by Brazilian presidents since 1985. Scholarly research about environmental policy in Brazil has focused on the role of the environmental bureaucracy [@silva-muller2022a], non-governmental organizations [@keck1998], legislation [@soares-filho2014], and international markets [@assuncao2015; @rajao2020 ]. Nevertheless, the role of the president remains understudied. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify the public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations of local actors, generated from material and discursive governmental practices, are a crucial factor in decisions to deforest on the ground [@assuncao2015; @capobianco2019; @capobianco2021; @campbell2015]. Second, we argue that presidents can deploy problem-constructions that build policy objects as different problems depending on whom they are speaking to. Presidential discourses take place in different settings, from launching a new bridge in a small municipality in the middle of the Amazon to a speech at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is discursively constructed in transnational politics.

To investigate how the Brazilian Amazon has been constructed as a problem in presidential speeches, we create a dataset containing 6130 official presidential speeches by all Brazilian presidents since 1985. We subset the dataset by identifying Amazon-related statements within these speeches. We find that 2014 sections in these discourses refer to the Amazon at least once. We then develop a codebook grounded in Amazonian historiography to code how each of these statements constructs Amazon as a particular problem. We use this codebook to manually code a randomly selected training set of Amazon-related statements. We train a supervised machine-learning model to automatically label the remaining set of Amazonian statements. We then conduct a descriptive and inferential analysis of this data, tying patterns in the data to Amazonian policies and deforestation outcomes over time.

Our empirical findings are threefold. First, we find that the frequency of the Amazon as a topic in speeches inside and outside Brazil is equally driven by domestic events (e.g. the assassination of Chico Mendes) as well as international events (e.g.Climate Summits). Second, the same government adopts a diverse range of problem-constructions over time. While economic integration dominated discourse from 1985 to the mid-2000s, environmental conservation and social development constructions steadily grew in the 2000s, briefly surpassing economic integration as more common problem-constructions from 2011 to 2016. Constructing the Amazon as an issue of sovereignty has become more pertinent after 2010. Finally, using logit models, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation as presidents speak farther away from the Amazon region. That is, presidents talk to the economic and social needs of the people domestically while boasting about environmental policy internationally.Empirically, we provide the first comprehensive overview of when, where, and how the Amazon is constructed as problems in presidential speeches. Conceptually, this contributes to understanding the social construction of the Amazon in discourses as a multi-level game and how these relates to policies.

This article proceeds as follows: first, we propose a conceptual framework to understand problem-construction in discourses. We then review Amazonian historiography literature to identify the main phases and their underlying problem-construction. In the methodology section, we explain how the framework is operationalized and present the codebook. We proceed to describe, visualize, and interpret our findings in the analysis section. Finally, we conclude with a discussion connecting our findings to past and present environmental policies.

# 2 Conceptual framework and contribution

## 2.1 Chosen problems, presidential discourses, and policy objects

In "Journey towards Progress", @hirschman1963 draws a conceptual distinction between pressing problems (pressured by outside parties to the government) and chosen problems (chosen by the government at their discretion). Pressing problems can be either privileged or neglected, depending on the degree of pressure exercised by the interested groups. Chosen problems are those that governments select at their discretion. Problems can change from pressing to chosen as a function of (a) solutions becoming available, (b) a change in the level of government control in society, or (c) a shift of interests from top policymakers [@hirschman1963]. @hirschman1975 example of policymakers choosing a problem is President's Kubitschek decision of building Brasilia, which was not particularly pressed by any interest group.

@bacchi2009 [p. 10] argues that policies have a cultural dimension as they take "shape within specific historical and national or international contexts". The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed, this is a policy object. We conceptualize policy objects as a problems that demand dedicated policy attention. The alleged problem is not always explicitly stated in a policy: policies are represented as solutions for implicit problems. Building Brasilia (the policy object), thus, could solve a problem of regional inequality, a problem of a dormant economy without state investment, a problem of political representation, or all three depending on how governments speak about it.

Governments can choose to emphasize (or not) one or more implicit problems that a policy solves depending on which interest group they are in communication with. Depending on how the policy is represented to be, it can be portrayed as a solution to problems that are considered pressing or not for different groups and it is up to the discretion of the political actor to construct a particular problem in a particular way given context. @putnam1988's seminal article on the two-level game can help us make sense of this variation conceptually. The author argues that the outcomes of international negotiations lie within the overlap between the agenda (and pressure) of domestic and international groups. If policies are socially constructed, governments can in theory construct them to meet the expectations of both. Consequently, this entails that governments are more diverse in their opinions, rather than monolithic.

An interesting context to study the relationship between problem-construction and policy objects is presidential speeches, as they influence ways of thinking about an issue. While the relationship between presidential speeches and policy may not be causal, they have the power to introduce and justify public policy, as well as shape issue perception to broad audiences [@zarefsky2004; @gillion2016]. As presidents speak across several locations to various audiences, they construct similar issues differently.

The possibility of different problem-constructions in presidential speeches has important implications for democracy and global politics. With globalization, presidents have the opportunity of promoting different problem-constructions transnationally. We understand transnationalism as relationships that transcend nation-states, from the local level to the international level encompassing non-state actors [@keck1998]. The environment and climate change, in particular, have received particular trasnantional attention, as narratives of global collective action dominated academic literature (for an overview, see @aklin2020).Relatedly, when Brazilian presidents speak about the Amazon it not only makes headlines, nationally and internationally [@brice2021; @harris2021; @miranda2021], but also incites responses, shapes expectations, and feeds into the behavior of a myriad of related actors. This is partially because the forest is a key biome in global climate-mitigation. It is important to understand, thus, to what extent the implemented policy agenda correlates with problem-constructions promoted at the local, federal, or international levels.

The connection between presidential discourse and the environment has been studied in the case of the United States but remains conceptually underdeveloped. @calderwood2019, for instance, examines 2019 mentions of climate change in American official presidential speeches since 1989. Among other things, the author finds that American presidents frequently side-step the environmental aspects of climate change. Building on @putnam1988, @calderwood2020 demonstrates that American presidents are more likely to mention climate change in foreign locations, and that location influences the specific discursive approach and tone they adopt. Elsewhere, @bevitori2015 finds that mentions of the environment are typically co-selected with the pronoun 'our', as well as with 'economy', 'clean', and 'preserve'. While these studies corroborate the possibilities of variation depending on the audience, they fail to tie vocabulary to specific problem-constructions based on wider shared meanings of environment and climate change. That is, problem-constructions touch on shared meanings available to both the speaker and the audiences as part of larger social-cultural history [see @bacchi2009], therefore, mentions to the lexicon as 'environment', 'preserve', or 'climate change should be tied to the larger meaning of policies in the United States.

There is also an empirical gap in terms of how the Amazon, specifically, has been constructed as a problem along time and geographic location in political discourses. While governmental discourses in Brazil have been studied for topics such as race relations [@silva2019] and inequality [@grangeia2017], the environment remains largely absent. An exception is @barros2020, who investigates Amazonian discourse in the Brazilian Congress. The author finds that the economic value of the Amazon for the cattle industry is the most salient narrative put forward by congressmen and concludes that there is a mismatch between the international debate (which focuses on preservation) and the national debate (which focuses on economic development).

This variation in problem-construction entails that the same governments are more diverse in their positions than the literature suggests. Scholars often describe federal governments as proponents of a cohesive set of policies toward the Amazon [see @drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. In this literature, the 1964 military dictatorship is associated with securing sovereignty in the region by populating it and integrating it into the national economy [@drummond2006]. The governments from the late 1980s up to 2009 are associated with a turn towards policies that focus on environmental conservation. The presidencies of Dilma and Temer (2011-2018) are connected to the decline of conservationist policies, while Bolsonaro to the dismantling of environmental policies. Although these works are important to understand how different governments have acted toward the Amazon, the classification of specific governments into policy periods, or policy cycles, represents them as monolithic: they associate specific governments with one specific view of the Amazon.

We propose a conceptual framework that accepts the possibility of varied problem-construction for the same policy object and connects it to presidential speeches. Governments often choose what problems to solve and what policies to implement. The same policies can be represented as solving different problems implying a degree of social construction [@bacchi2009]. Governments can emphasize, or not, implicit problems that a policy solves depending on which interest group they are in communication with. The variation in problem-construction suggests that Brazilian governments are more diverse in their positions than the literature suggests. An interesting context to study the relationship between varied problem-construction and policy is presidential speeches as they legitimize ways of thinking about an issue, matter for policymaking, and environmental outcomes. As presidents speak in different places and to diverse audiences, it is an empirical site prone to identifying variation.

## 2.2 Problem-construction in Amazonian historiography

We rely on the Amazonian historiography literature to identify possible problem-constructions and their connection to the socio-cultural history of the country. We refer to Amazonian historiography as the body of research conducted by social and environmental scientists that tell the story of diverse governmental programs that approach the Amazon region, forest, and its people as a policy object demanding dedicated action. From this literature, we identify three historical-political constructions: national sovereignty, economic integration, and environmental conservation. We also argue a fourth, social development, that focuses on the rights and dignity of the inhabitants of the Amazonian region.

### National Sovereignty

@hecht1990 [p. 1] write that "what imbues the case of the Amazon with such passion is the symbolic content of the dreams it ignites". In the process of securing Amazonian borders since the 18th century, Brazil thwarted "the imperial ambitions of France, Britain, the United States, Belgium, Bolivia, and Peru" [@hecht1990, p. 8], and when the dust settled and the scramble was over, over half of the Amazon emerged Brazilian. While Brazilian military diplomacy was very successful, the process did not come without its traumas as for example the negotiations with Bolivia in 1902 to secure the territory of the state of Acre [@hecht1990]. These traumas were referred to, and offered context, to Amazonian discourses and policies during the military dictatorship of 1964 as earlier regimes aimed at protecting Brazil's sovereignty over the Amazon [@hecht2020]. As we move to a world where non-state actors gain importance in environmental governance [@silva-muller2022; @andonova2014; @keck1998], sovereignty-related problems become more varied. Threats to national sovereignty, consequently, broaden from nation-states to a wide set of actors. Claims of attempts to 'internationalize' the Brazilian Amazon, for instance, are targeted at foreign actors as well as domestic non-state actors. The sovereignty problem-construction advances the view that the Brazilian Amazon belongs to Brazil and any foreign or non-state presence in the region can be part of a broad strategy to take the region. The policy solutions to the issue of sovereignty included the monitoring of the borders and strict regimes related to entry into the region.

### Economic Integration

The Vargas dictatorship (1937-46) and the military dictatorship (1964-89) took over the task of modernizing the Amazon. In 1966, the Brazilian Military launched 'Operation Amazon', a policy to modernize the region based on three assumptions[@acker2021] . First, it assumes that nature should be conquered by mankind Second, it assumes that exploiting natural resources would render the Amazon region economically profitable. Third, it assumes that populating the region was necessary to integrate it into the country and exert control over the territory. Concretely, this meant a series of major infrastructural projects, incentives for settlers to expand the agricultural frontier, and the establishment of a tax-free zone in Manaus to attract industry. @capobianco2019 and @becker2005 describe the period from the 1950s to the 1980s in similar vein, as one of often centralized economic policies by the federal government.The economic integration problem-construction advances the view that the Brazilian Amazon needs to be developed and modernized. These policy solutions have at their core the development of the necessary infrastructure (physical, fiscal, or monetary) to integrate the region into the national and international economy.

### Environmental Conservation

The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of environmental institutions [@drummond2006]. A common explanation for the creation of these institutions in the Amazonian literature is the impression of the lack of control over the market, engendered by years of centralized economic integration in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987], epitomized in the 1992 Earth Summit in Rio de Janeiro [@hochstetler2021; @capobianco2021]. @hochstetler2007 argue that during preparations for the summit, a new form of Brazilian environmentalism emerged: socio-environmentalism. They define it as an emphasis on the local livelihoods of traditional populations while protecting nature. @capobianco2019 argues that socio-environmentalism informed a series of policies throughout the 1990s and early 2000s towards preserving the forest, such as the establishment of several conservation units in the Amazon in 2001. The environmental conservation problem-constructions emphasize that the Amazon should be preserved, deforestation should be halted, and the sustainable practices of indigenous and local peoples should be maintained through the protection of their territories and rights to self-determination [@hochstetler2007]. The policy solution implies, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

### Social Development

Governments can emphasize the lack of hospitals, sanitation, and schools concerning peoples' dignity, standards of living, and other constitutional rights. However, such emphasis can be masked within both the economic integration and environmental conservation accounts. @acker2021 [p. 10], for example, argues that during the military dictatorship attempts to populate the Amazon were seen, among other things, as a solution for the country's growing social inequalities elsewhere. As democracy established itself in Brazil, the settlers from the military dictatorship became electorates with constitutional rights. @hochstetler2007 argue that with the transition to democracy, the mobilization of diverse social movements to demand social equality and conservation was an integral part of the socio-environmentalist movement in Brazil. While indigenous rights and preservation of the forest go hand in hand, it is not clear whether providing better standards of livings to the large and poor populations of the state and preservation do; perhaps this is also why there was never an open consensus about socio-environmentalism in Brazil [@hochstetler2007]. Although related to economic integration and environmental conservation, social development is a problem-construction that focuses on constitutional rights, citizenship, and dignity. The policy solution to social development implies the policies to facilitate access to water, sanitation, electricity, internet, and radio, as well as the construction of schools and hospitals locally.

# 3 Methodology: operationalizing Amazonian problem-constructions

## 3.1 Codebook and Data

To analyze how presidents construct the Amazon as a problem, we build upon the dataset provided by [@cezar2020] which contains all official speeches by Brazilian Presidents from 1985 to 2019 scrapped from the archives of the Brazilian Presidential Library. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches for all the presidents of Brazil. Then, we proceed to identify all speeches that refer to the Amazon as a region, people, or forest. We do so by detecting all speeches in which the stem "amazon" appears. In Portuguese, the stem captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We find that 946 speeches are, at least partially, about the Amazon out of the 6130.

Using the poldis R package [@sposito2021], we proceed to extract two sentences before and two sentences after the sentence in which the stem "amazon" appears. We opt for picking two sentences around, rather than words, because sentences usually contain a cohesive idea. By doing so we create our unit of analysis: an Amazonian statement. We use Amazonian statements as our unit of analysis because it allows us to identify only passages that are meaningful for our specific purpose.This process yields 2014 unique Amazonian statements about the Amazon. When an Amazonian statement contains two or more matches of the stem "amazon", we get two sentences before the first match and two sentences after the last match. On average, an Amazonian statement contains 123 words.

Our approach to text coding has three steps. First, we develop a codebook to code Amazonian statements in one or more problem-constructions (see codebook in appendix). In their conceptualization, each problem-construction is mutually exclusive, meaning that they cover different forms of constructing the Amazon as a problem. Nevertheless, each Amazonian statement might be assigned to one or more codes. A statement can, for example, construct the Amazon as a problem of sovereignty and a problem of economic integration, or a problem of social development and conservation. Amazonian statements, thus, can be either coded as pure-types or mixed-types. Mixed-types are relevant as constructions and policies are often multifaceted while portraying the Amazon as a combination of these problem-constructions. Figure 1, below, portrays this operationalization strategy.

@@

Second, with the codebook in hand, each one of the authors, separately, hand-coded the same set of 1007 randomly selected Amazonian statements. This amount refers to 50% of all the Amazonian Statements identified. We chose to hand code half of the observations because there are several nuances in discourse in how presidents talk about the Amazon over time and as a problem. As the size of the training set should increase with the number of categories [@grimmer2022], we deemed four categories and 50% of the statements as reliable. This allows a robust validation set to verify the models. Automating the coding of half of the observations saved the authors over one month of work in comparison to manual coding. The intercoder agreement for hand-coded categories was 85%, on average. For each non-matching coded observation, the co-authors discussed and sorted their disagreements, which were subsequently specified in the codebook.

Third, the hand-coded data is then randomly divided into a training set, containing 80% of the hand-coded observations (806 observations), and a validation set, containing the remaining 20% of the hand-coded data (201 observations). We chose to employ a support-vector machine (SVM) algorithm, a non-probabilistic linear classifier that classifies documents by assigning points in mapped space to maximize the gap between binary categories, to label texts [@meyer2021; @noble2006]. The SVM model is trained using the hand-coded training set and then employed to classify observations in the validation set. The trained SVM model was, on average 82%, accurate in classifying observations in the validation set before being tuned. After the SVM model is tuned, we use the model to automatically code the remaining 1007 Amazonian statements. The final dataset for analysis, excluding false positive matches, contains 1895 coded Amazonian statements.

@@

To analyze our data, we first present a series of different plots on the proportions of Amazonian statements and problem-constructions over time and by presidents. To test whether problem-constructions depend on where presidents speak, we run four separate logistic regressions. Each model takes the share of a specific problem-construction versus all other problem-constructions as their dependent variable. This means we predict the probability of a president constructing the Amazon as an issue of (1) environmental conservation, (2) economic integration, (3) social development, and (4) national sovereignty, separately, and in relation to the remaining three problem-constructions. Our main independent variable is the distance (in kilometers) from the state or country capital where the speech took place to Manaus (the capital of the Amazonas State and centrally located in the region). The logic behind this relates to presidents playing multi-level games. We expect presidents to change how they construct the Amazon based on their believes of audiences' expectations.We interpret the plots and model considering multiple Amazon-related events and policies over the last 30 years, with the support of existing literature, and their correlations with different presidents and locations.

This approach comes with limitations. Our codebook is developed using specific Amazon-related vocabulary. For example, a statement will be coded as economic integration if it is meaningful support for the tax-free zone of Manaus or a dam in the Amazon. However, the economy is generally a topic that presidents speak about frequently. Hence, the high incidence of economic integration in Amazonian statements can also be related to the higher importance of this problem-construction in Brazil overall. Moreover, we classify statements as Amazonian based on a dictionary composed of a single lexicon stem: "amazon". We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon "Amazon", for example, when the president says, "the forest" or "deforestation". Hence, we might be missing statements about Amazon that do not refer to it. However, we consider this safer as we cannot be sure that mentions of the forest or deforestation do not correspond to other biomes such as the Cerrado or the Atlantic Forest. Finally, our dataset covers only what is considered an official speech. Presidents, though, give interviews, appear in debates, talk at campaign rallies, and, more recently, post on social media. Problem-construction within presidential discourse, thus, also happens in different sites for which we do not account in this paper.

# 4 How has the Amazon been constructed as a problem?

## 4.1 The rises and falls of the Amazon as a topic in presidential speeches

When, where, and how often do presidents speak about the Amazon? Figure 2, below, shows the proportion of speeches that mentions the Amazon in relation to all speeches in a year. On average, presidents mention the Amazon in 15.4% of their speeches. For comparison purposes the averages for other policy objects in the same corpus of speeches are: inequality (14.7%), criminality (17.3%), inflation (20.9%), and unemployment (13.8%). The frequency at which the Amazon appears across presidential speeches inside, outside, and overall varies similarly across time (see Figure 7 in appendix).

```{r Figure 2, fig.cap="Amazonian speeches in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE}
perc1 <- pop %>%
  count(amazon_speech, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>% 
  select(c(perc, amazon_speech, year))
Amazon_year <- perc1 %>% ggplot(., aes(x = year, y = perc)) +
  geom_line(color = "darkgreen") +
  geom_line(aes(y=mean(perc)), color="black", size=.5, linetype="dotted")+
  scale_y_continuous(labels = percent_format()) + 
  annotate("text", x = 1985, y = 0.18, label = "15.4%", size = 3.5)+
  annotate("text", x = 1988.7, y = 0.35, label = "The Murder of\nChico Mendes", size = 3)+
  annotate("text", x = 1992.5, y = 0.27, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.24, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2015, y = 0.21, label = "Paris Summit", size = 3)+
  annotate("text", x = 2019, y = 0.26, label = "2019 Record\nAugust Fires", size=3) +
  labs(x = "", y = "",
       title = "Share of speeches that mention the Amazon per year",
       subtitle = "Sample composed by all 6130 speeches since 1985",
       color='Y series') +
  theme_am()
Amazon_year
```

Five local maxima stand out in the figure: 1989, 1992, 2009, 2015, and 2019. These points coincide with internal and external events that help us explain the rises and falls of the Amazon in presidential discourse. From 1985 to 1988, the mentions of the Amazon in official speeches increased from 4% to 14%. During this period, the latest Brazilian Constitution was written, which gave indigenous and traditional populations space to advocate for constitutional environmental rights and the protection of their territories [@hecht1990]. In 1989, however, the Amazon appeared in 32% of all speeches. This maximum correlates with the brutal murder of Chico Mendes in the last days of 1988. The incident caught unprecedented international attention and Sarney responded to this with a set of policies to address deforestation [@capobianco2021], including accepting to host the 1992 Earth Summit [@keck1998].

1990-91 the Amazon appeared in 12% to 14% of all speeches but, by 1992, this average increased to 25%. The 1992 Earth Summit in Brazil brought international attention to environmental topics related to the Amazon. One of the big announcements, for example, was the consolidation of the first transnational partnership for the Amazon, the G7 Pilot Program, which brought a high number of financial resources to the region for public policy implementation [@capobianco2021]. Throughout the rest of the 1990s and 2000s, during the Cardoso (1995-2002) and Lula administrations, the Amazon as a topic did not diverge much from the average until 2009, when we see another maximum in mentions of the Amazon. This coincides with the 2009 Copenhagen Summit and the steepest decrease in deforestation rates since data is available as Figure 3, below, illustrates. Lula led the delegation to Copenhagen with a self-image of "we do not promise, we deliver" when the stakes about climate change were high [@franchini2019].

```{r Figure 3, fig.cap="Deforestation rates since 1988", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
def_rates <- ggplot(amazon_def_year, aes(x = year)) +
  geom_line(aes(y = area), color= "darkolivegreen", linetype="twodash", size= 0.5)+
  geom_smooth(se=FALSE, color="black", aes(y=area), linetype="solid", size= 0.5)+
  annotate("text", x = 2005, y = 29000,
           label = "2004 Plan to Combat and Control Deforestation", size = 3) +
  annotate("text", x = 2012.5, y = 3500, label = "2012 Forest Code ", size = 3) +
  labs(x = "", y = "Deforestation in squared kilometers",
       title = "Deforestation rates since 1988",
       subtitle= "Data from INPE, Brazilian Ministry of Science and Technology.",
       color = "",
       caption= "Black curve smoothed using loess.")+
  theme_am()
def_rates
```

>>>>

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(readxl)
library(ggplot2)
library(gridExtra)
library(modelsummary)
library(plm)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

@@

        peak_year = ifelse(grepl("1989|1992|2009|2015|2019", year), 1, 0),
        py_1989=ifelse(grepl("1989", year), 1, 0),
        py_1992=ifelse(grepl("1992", year), 1, 0),
        py_2009=ifelse(grepl("2009", year), 1, 0),
        py_2015=ifelse(grepl("2015", year), 1, 0),
        py_2019=ifelse(grepl("2019", year), 1, 0),

@@

                                            location), "Amazonian Countries",
                                      location_cat),
                               levels = c("Amazonian States",
                                          "Non Amazonian States",
                                          "Brasilia",
                                          "Amazonian Countries",
                                          "International")))

ama_model_lag <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") %>%
dplyr::left_join(AAI, by = "year") %>%
filter(location_cat != "Non Identified", mx_cat != "Other Mixed-types") %>%
  mutate(con_vs_all = ifelse(mx_cat == "Pure Environmental Conservation", 1, 0),
         EI_vs_all = ifelse(mx_cat == "Pure Economic Integration", 1, 0),
         SD_vs_all = ifelse(mx_cat == "Pure Social Development", 1, 0),
         sov_vs_all = ifelse(mx_cat == "Pure National Sovereignty", 1, 0))
```

@@

# 3 Methods

## 3.1 Data and Operationalization

To analyze how presidents construct the Amazon as a problem, we build upon the dataset provided by (Cezar 2020) which contains all official speeches by Brazilian Presidents from 1985 to 2019 scrapped from the archives of the Brazilian Presidential Library. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches. Then, we proceed to identify speeches that refer to the Amazon as a region, people, or forest. We do so by detecting all speeches in which the stem “amazon” appears. In Portuguese, the stem captures terms such as “Amazonia”, “Amazonica”, “Amazonidas”, “Amazonense(s)”, “Amazonas”, among others. We find that 946 speeches are, at least partially, about the Amazon.

Using the poldis R package (Sposito 2021), we proceed to extract two sentences before and two sentences after the sentence in which the stem “amazon” appears. We opt for picking two sentences around, rather than words, because sentences usually contain a cohesive idea. When an Amazonian statement contains two or more matches of the stem “amazon”, we get two sentences before the first match and two sentences after the last match. By doing so we create our unit of analysis: an Amazonian statement. Amazonian statements represent text portions that are meaningful for our specific purpose. This process yields 2014 unique Amazonian statements about the Amazon. On average, an Amazonian statement contains 123 words.

Our approach to text coding is threefold. First, we follow the Amazonian historiography literature to develop a codebook to code Amazonian statements as one, or more, problem-constructions (see codebook in appendix). In their conceptualization, each problem-construction is mutually exclusive, meaning that they cover different forms of constructing the Amazon as a problem. Nonetheless, each Amazonian statement might be assigned to one or more codes. A statement can, for example, construct the Amazon as a problem of sovereignty and a problem of economic integration, or a problem of social development and conservation. Amazonian statements, thus, can be either coded as pure-types or mixed-types. Figure 1, below, portrays this operationalization strategy.

@@

Second, with the codebook in hand, each one of the authors, separately, hand-coded the same set of 1007 randomly selected Amazonian statements. This amount refers to 50% of all the Amazonian Statements identified. We chose to hand code half of the observations because there are several nuances in discourse in how presidents talk about the Amazon over time, and as a problem. As well, the size of the training set should increase with the number of categories (Grimmer, Roberts, and Stewart 2022). Since we have four categories, hand coding 50% of the statements mean we have a robust and reliable number of observations to be used for the training, and for validation, of the models. Automating the coding of half of the observations saved the authors over one month of work in comparison to manual coding. Once coded, the average intercoder agreement for hand-coded categories was 85%. For each non-matching coded observation, the authors discussed and sorted their disagreements, which were subsequently specified in the codebook.

Third, the hand-coded data is then randomly divided into a training set, containing 80% of the hand-coded observations (806 observations), and a validation set, containing the remaining 20% of the hand-coded data (201 observations). We chose to employ a support-vector machine (SVM) algorithm, a non-probabilistic linear classifier that classifies documents by assigning points in mapped space to maximize the gap between binary categories, to label texts (Meyer et al. 2021; Noble 2006). As a robustness check, the SVM model is trained using the hand-coded training set and then employed to classify observations in the validation set. The trained SVM model was, on average, 82%, accurate in classifying observations in the validation set before being tuned. After the SVM model is tuned and the size of the training set is increased by including all hand coded observations, we use the model to automatically code the remaining 1007 Amazonian statements. The authors then randomly select 100 Amazonian statements automatically coded and find that the model is 97% accurate in coding these statements. The final dataset for analysis, excluding false positive matches, contains 1895 coded Amazonian statements (add a footnote for false positives here)).

@@

To analyze our data, we model the share of a specific pure problem-construction, versus all other problem-constructions, as the dependent variable, in relative to the settings in which a particular speech happens, as our independent variable. This means we predict the probability of a president constructing the Amazon as an issue of (1) environmental conservation, (2) economic integration, (3) social development, and (4) national sovereignty, separately, and in relation to the remaining problem-constructions (see table 1 below). We chose to focus on pure constructions to provide unambiguous interpretation and robust correlations. We divide setting in terms of non-Amazonian Brazilian states, Amazonian states within Brazil, Brasilia, Amazonian countries, and international. The level aggregation refers to geo-political settings that make sense in light of the Amazon, as a policy object. This aggregation combines different expectations presidents might have about transnational audiences in these settings. As well, Brasilia is a separate setting because of its’ unique location and since most presidential speeches are delivered in Brasilia. We control for yearly deforestation rates, inflation, election years, and peak years for Amazon appearing year in presidential speeches. Subsequently, we plot Amazonian statements and problem-constructions over time and by setting. We interpret the plots and model considering the literature alongside with multiple Amazon-related events and policies over the last 30 years.

```{r table 1, fig.cap="Models Description", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE}

models <- tibble::tibble("Models Description" = c("Fixed-effects logistic regressions indexed by president (controlling for time and ideology)."),
                         "Dependent variables" = c("Pure environmental conservation; pure economic integration; pure social development; and pure national sovereignty."),
                         "Independent Variables" = c("Amazonian states within Brazil (reference category); non-Amazonian Brazilian states; Brasilia; Amazonian countries; and international."),
                         "Controls" = c("Yearly deforestation rates lagged by one year; average annual inflation; election years; and peak years for Amazon appearing year in presidential speeches (footnote)."))

kbl(models, booktabs = T, caption = "Model description") %>%
  kable_styling(latex_options = c("striped", "repeat_header",
                                  "scale_down", "HOLD_position"),
                position = "center", font_size = 9, full_width = TRUE)
```

The main limitations of our approach are threefold. First, related to our sub-setting and coding strategies, we classify statements as Amazonian based on a dictionary composed of a single lexicon stem: “amazon”. We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon “Amazon”, for example, when the president says, “the forest” or “deforestation”. Hence, we might be missing statements about Amazon that do not refer to it. However, we consider this safer as we cannot be sure that mentions of the forest or deforestation do not correspond to other biomes such as the Cerrado or the Atlantic Forest. Moreover, our codebook is developed using specific Amazon related vocabulary. For example, a statement will be coded as economic integration if it is meaningful support for the tax-free zone of Manaus or a dam in the Amazon. However, the economy is generally a topic that presidents speak about frequently. Hence, the high incidence of economic integration in Amazonian statements could also be related to the higher importance of this problem-construction in Brazil overall.

Second, our dataset covers only what is considered an official speech. Presidents, though, give interviews, appear in debates, talk at campaign rallies, and, more recently, post on social media. Problem-construction within presidential discourse, thus, happens in different public and private venues for which we do not account in this paper. We believe that setting remains an important factor in changing expectations about audiences, especially for a historical analysis. Although we recognize that this, arguably, has begun to change with the advancement of communication technologies. Furthermore, we aggregate levels to socio-political-cultural settings that make sense in light of the policy object in question. Therefore, we are aware that a category as “international”, for instance, encompasses speeches delivered by presidents in a wide variety countries and institutions to a wider variety of audiences, which are not differentiated.

Third, employing lagged deforestation as a control variable has several constraints. We assume presidential discourse, policies, and outcomes are co-constitutive. Though, we lag annual deforestation one year since data for the year usually gets publicly released later and, frequently, there is a time difference to when discourses might refer to certain deforestation rates. We do think discourses capture general problem constructions that can be used to introduce, discuss, and justify policy and general outcomes. Throughout the analysis we carefully discuss how presidential discourses might relate to deforestation rates and outcomes by pointing to the patterns and correlations without making causality claims.
